Algorithms for Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2013-10-19 paper/paper.tex

Large-scale scientific workflows, usually represented as Directed Acyclic Graphs
(DAGs), are an important class of applications that lead to challenging problems
in resource management on grid and cloud infrastructures. What becomes more
important, workflows for large computational problems are often composed of
several inter-related workflows grouped into {\em ensembles}.  Workflows in an
ensemble typically have a similar structure, but they differ in their input
data, number of tasks, and individual task sizes.

There are many applications that require scientific workflow ensembles.
CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate seismic
hazard maps.  Each workflow in a CyberShake ensemble generates a hazard curve
for a particular geographic location, and several hazard curves are combined to
create a hazard map.  In the recent study of 2013 CyberShake generates a map that
requires an ensemble of 286 workflows.  Similarly, users of
Montage~\cite{Deelman2008} often need several workflows  with different
parameters to generate a set of image mosaics that can be combined  into a
single, large mosaic. The Galactic Plane ensemble, which generates several
mosaics of the  entire sky in different wavelengths, consists of 17 workflows,
each of which contains 900 sub-workflows.  Another ensemble example is the
Periodograms application~\cite{Vockler2011}, which searches for extrasolar
planets by detecting periodic dips in the light intensity of their host star.
Due to the large scale of the input data, this application is often split up
into multiple batches processed by different workflows. Additional workflows are
created to run the analysis using different parameters. A recent analysis of
Kepler satellite data required three ensembles of 15 workflows.

Workflows in an ensemble may differ not only in their parameters, but also in
their priority. For example, in CyberShake some geographic locations may be in heavily
populated areas or in strategic locations such as power plants, while others
may be less important. Scientists typically prioritize the workflows in such
an ensemble so that important workflows are finished first. This enables them
to see critical results early, and helps them to choose the most important
workflows when the time and financial resources available for computing are
limited.

@@

The granularity of the workflow tasks should be discussed in the context of the
length of the cloud billing cycle. In this paper we assumed that the resources
are charged for hourly time slots, as is the case of VM instances on Amazon EC2.
However, there are cloud providers that have higher billing frequency, e.g.
every 5 minutes in the case of
CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1 minute in the
case of Google Compute
Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
first 10 minutes). Our results indicate that when the task granularity is close
to the length of the billing period, the SPSS produces better results, while for
tasks that are shorter than the billing period, the dynamic algorithms perform
better. This suggests that these granularity effects have to be taken into account when
selecting a cloud provider, and also when using task clustering algorithms that
group smaller tasks into larger groups~\cite{ChenD12}.

@@

Figure~\ref{fig:variances-scores} shows relative performance as a percentage of
the number of high scores achieved by each algorithm as the error increases. As
in Figure~\ref{fig:completed-budgets} and ~\ref{fig:completed-deadlines} we show
results for CyberShake and LIGO as representatives of fine grained and coarse
grained workflows. The figure shows that, for this experiment, the scores remain
the same as the error increases. This is a result of the way the error is
applied. Since the error is selected uniformly from $-p$ to $+p$ it is equally
likely that the error will be positive as negative. Therefore, applying the
error leaves the total runtime of the tasks in the workflow unchanged -- the
total number of CPU hours in each workflow remains nearly the same regardless of
the error. The end result is that the dynamic algorithms are able to achieve the
same score without exceeding the constraints by using a different schedule that
finishes the same total amount of work using the same budget and deadline.

@@

In cases where there are no provisioning delays, task runtime estimates are
good, and failures are rare, we found that SPSS performs significantly better
than both dynamic algorithms. However, when conditions are less than perfect,
the static plans produced by SPSS are disrupted and it frequently exceeds the
budget and deadline constraints. In comparison, the dynamic algorithms are able
to adapt to a wide variety of conditions, and rarely exceed the constraints even
with long delays, poor estimates, and high failure rates. This, however, comes
at the cost of being able to complete less workflows than SPSS. Alternatively, a
user may decide to accept that the budget and deadline constraints are exceeded,
as long as more work is completed.

>>>>

Large-scale scientific workflows, usually represented as Directed Acyclic Graphs
(DAGs), are an important class of applications that lead to challenging problems
in resource management on grid and cloud infrastructures. In addition, workflows
for large computational problems are often composed of several inter-related
workflows grouped into {\em ensembles}. Workflows in an ensemble typically have
a similar structure, but they differ in their input data, number of tasks, and
individual task sizes.

There are many applications that require scientific workflow ensembles.
CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate seismic
hazard maps.  Each workflow in a CyberShake ensemble generates a hazard curve
for a particular geographic location, and several hazard curves are combined to
create a hazard map.  In a 2013 study, CyberShake was used to generate a set
of hazard maps over 286 sites that required an ensemble of 2288
workflows~\cite{CyberShake13.4}. Similarly, users of Montage~\cite{Deelman2008}
often need several workflows  with different parameters to generate a set of
image mosaics that can be combined  into a single, large mosaic. The Galactic
Plane ensemble~\cite{GalacticPlane}, which generates several mosaics of the
entire sky in different wavelengths, consists of 17 workflows,
each of which contains 900 sub-workflows.  Another ensemble example is the
Periodograms application~\cite{Vockler2011}, which searches for extrasolar
planets by detecting periodic dips in the light intensity of their host star.
Due to the large scale of the input data, this application is often split up
into multiple batches processed by different workflows. Additional workflows are
created to run the analysis using different parameters. A recent analysis of
Kepler satellite data required three ensembles of 15 workflows.

Workflows in an ensemble may differ not only in their parameters, but also in
their priority. For example, in CyberShake some geographic locations may be in heavily
populated areas or have sensitive facilities such as power plants, while others
may be less important. Scientists typically prioritize the workflows in such
an ensemble so that important workflows are finished first. This enables them
to see critical results early, and helps them to choose the most important
workflows when the time and financial resources available for computing are
limited.

@@

The granularity of the workflow tasks should be discussed in relation to the
length of the cloud billing cycle. In this paper we assumed that the resources
are billed by the hour, as is the case of VM instances on Amazon EC2.
However, there are cloud providers that have higher billing frequency, e.g.
every 5 minutes in the case of
CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1 minute in the
case of Google Compute
Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
first 10 minutes). Our results indicate that when the task granularity is close
to the length of the billing period, the SPSS produces better results, while for
tasks that are shorter than the billing period, the dynamic algorithms perform
better. This suggests that these granularity effects have to be taken into account when
selecting a cloud provider, and also when using task clustering algorithms that
group smaller tasks into larger groups~\cite{ChenD12}.

@@

Figure~\ref{fig:variances-scores} shows relative performance as a percentage of
the number of high scores achieved by each algorithm as the error increases. As
in Figure~\ref{fig:completed-budgets} and ~\ref{fig:completed-deadlines} we show
results for CyberShake and LIGO as representatives of fine-grained and
coarse-grained workflows. The figure shows that, for this experiment, the scores remain
the same as the error increases. This is a result of the way the error is
applied. Since the error is selected uniformly from $-p$ to $+p$ it is equally
likely that the error will be positive as negative. Therefore, applying the
error leaves the total runtime of the tasks in the workflow unchanged -- the
total number of CPU hours in each workflow remains nearly the same regardless of
the error. The end result is that the dynamic algorithms are able to achieve the
same score without exceeding the constraints by using a different schedule that
finishes the same total amount of work using the same budget and deadline.

@@

In cases where there are no provisioning delays, task runtime estimates are
good, and failures are rare, we found that SPSS performs significantly better
than both dynamic algorithms. However, when conditions are less than perfect,
the static plans produced by SPSS are disrupted and it frequently exceeds the
budget and deadline constraints. In comparison, the dynamic algorithms are able
to adapt to a wide variety of conditions, and rarely exceed the constraints even
with long delays, poor estimates, and high failure rates. However, this comes
at the cost of not being able to complete as many workflows as SPSS in ideal
conditions.
